---
title: "SPADI example"
author: "Andreas Kryger Jensen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SPADI example}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
```


Optionally install the R package
```{r, eval = FALSE}
remotes::install_github("ERRTG/gRm")
```

Then we load the package.
```{r}
library(gRm)
```

Load the data and inspect the names of the variables
```{r}
SPADI <- read.csv(url("https://erda.ku.dk/public/archives/bacc560d26b01f7a65e77a9712a92e86/SPADI.csv"))
# The source file contains one isolated P1 = 0 observation whose remaining
# record state leaves that category outside the conditional estimation sample.
# Remove that unusable category before inferring the vignette's item levels so
# every declared category has source estimation support.
SPADI <- SPADI[is.na(SPADI$P1) | SPADI$P1 > 0, , drop = FALSE]
colnames(SPADI)
```


# Defining an analysis

We start by defining a gRm analysis object based on five items and two exogenous variables from the data set. Calling the resulting object implicitly executes the print function of the object which gives basic information. For more information you can call summary on the object. Score groups are automatically defined by a median split, but they can also be specified explicitly by an argument in the construction.

```{r}
spadi_proj <- gRm(SPADI, 
                  items = sprintf("P%01d", 1:5),
                  exogenous = c("over60", "gender"))
spadi_proj
summary(spadi_proj)
```

# Rasch model

The next step is to set up a GLLRM model. We will start by fitting a standard Rasch model, so the call to gllrm is very simple. Calling the resulting object gives you basic information about the model. We see that there are neither LD nor DIF in the model. We will later see how to specify these and thus extend the Rasch model to a Graphical Log-Linear Rasch Model.

```{r}
m_rasch <- gllrm(spadi_proj)
m_rasch
```

Then we can fit the model. 

```{r}
fit_rasch <- fit(m_rasch)
fit_rasch
```

The print function used above gives a condensed summary, while the summary function gives more information. 

```{r}
summary(fit_rasch)
```

It contains two main blocks, item parameters and thresholds. The plain summary command prints both but they can also be individually printed by using the which argument.

```{r}
summary(fit_rasch, which = "parameters")
```

## Item fit

Item fit has two different views. The default is the tests view,

```{r}
item_fit(fit_rasch) #defaults to which = "tests"
```

Additional information is given by setting the view equal to items.

```{r}
item_fit(fit_rasch, which = "items")
```


## Test of global homogeneity

Test for global homogeneity follows a similar pattern where print gives the brief summary and the summary function exposes more information. The global homogeneity function will by default use the score cuts defined when constructing the analysis object, but different cuts can explicitly be defined when calling the function thus overriding the already defined ones.


```{r}
gh_rasch <- global_homogeneity(fit_rasch)
gh_rasch
```

```{r}
summary(gh_rasch)
```

## Check for local dependencies

We can check for local dependence using the local_dependence function.

```{r}
ld_rasch <- local_dependence(fit_rasch)
ld_rasch
```

```{r}
summary(ld_rasch)
```

## Check for DIF

Similarly, we can check for DIF using the dif function.

```{r}
dif_rasch <- dif(fit_rasch)
dif_rasch
```

```{r}
summary(dif_rasch)
```


# Screening

The package also implements the SCREEN J procedure from DIGRAM.

```{r}
spadi_screen <- screen(spadi_proj)
spadi_screen
```

```{r}
summary(spadi_screen)
```

The result from the screening can automatically be turned into a GLLRM model
by feeding it into `gllrm()`. The resulting model contains the LD and DIF terms
identified by SCREEN J.

```{r}
m_screen <- gllrm(spadi_screen)
m_screen
```

A visualization of the model can be obtained by the plot function.

```{r, fig.width=6, fig.height=4, out.width="100%", dpi=192}
plot(m_screen)
```

SCREEN J proposes candidate terms; the public fit boundary still verifies the
source item-category and interaction-margin support before estimation. For
these data the complete screened model has the required support and converges.

```{r}
screen_fit <- fit(m_screen)
screen_fit
```

For comparison, we also demonstrate a smaller custom GLLRM chosen from the
same candidate families.

# Setting up a general GLLRM

A custom GLLRM can also be defined by giving the specific LDs and DIFs as one-sided formula objects.

```{r}
m_gllrm <- gllrm(spadi_proj, ld = ~ P3:P4, dif = ~ P4:over60)
```

We can do a direct likelihood ratio test of the GLLRM against the Rasch model

```{r}
fit_gllrm <- fit(m_gllrm)
anova(fit_rasch, fit_gllrm)
```

Global homogeneity for the fitted GLLRM includes the applicable uniform local
dependence and DIF sections.

```{r}
summary(global_homogeneity(fit_gllrm))
```


# ARI plot

The package also includes ARI plots

```{r, fig.width=6, fig.height=4, out.width="100%", dpi=192}
fit_gllrm |> ari() |> plot()
```

# CM2

The following CM2 and CM3 examples use two focal-item scopes. Omitting `items`
selects all five fitted items, which is the R equivalent of accepting DIGRAM's
blank item-selection prompt. We then repeat each diagnostic with the explicit
three-item subset `P1`, `P3`, and `P5`.

The subset changes only the margins being tested. The probabilities still come
from the complete fitted GLLRM: the unselected items `P2` and `P4`, the fitted
`P3:P4` local-dependence term, the fitted `P4:over60` DIF term, both exogenous
variables, and the total score across all five fitted items remain part of the
computation.

The compact result and full `summary()` printout use the same display precision
as the package's other diagnostic summaries. The tables returned by
`summary(x)$tables` remain unrounded numeric data; printing one of those tables
directly, as the aggregate examples below do, delegates its display precision
to base R. Empirical bootstrap zero is printed as exact `0`, and a bootstrap
seed, when shown in the full summary, is printed as its exact unsigned decimal
identifier.

```{r}
diagnostic_subset_items <- c("P1", "P3", "P5")
```

## Asymptotic CM2

The ordinary asymptotic diagnostic is the default (`bootstrap = FALSE`). The
first call deliberately omits `items`, so every fitted item is focal.

```{r}
cm2_all_asymptotic <- cm2(fit_gllrm)
cm2_all_asymptotic
summary(cm2_all_asymptotic)$tables$aggregates
summary(cm2_all_asymptotic)
```

The second call changes only the focal margin set. Its selection metadata
records the requested names and their source-ordered resolution.

```{r}
cm2_subset_asymptotic <- cm2(
  fit_gllrm,
  items = diagnostic_subset_items
)
cm2_subset_asymptotic
cm2_subset_asymptotic$metadata$selection$resolved_items
summary(cm2_subset_asymptotic)$tables$aggregates
summary(cm2_subset_asymptotic)
```

## Bootstrap CM2

Setting `bootstrap = TRUE` requests DIGRAM's source-shaped parametric
bootstrap. The same fixed seed is used for the all-item and subset calls. Since
selection affects only the tested margins, the two calls generate and refit the
same full-model samples; they differ in the CM2 statistics evaluated on those
samples.

The vignette uses only 10 samples per call to keep its build time reasonable.
This demonstrates the interface but is too small for substantive inference. A
real analysis should request a suitably large value such as `nsim = 1000L` and
inspect `nused` before interpreting bootstrap p-values.

```{r}
vignette_bootstrap_nsim <- 10L
vignette_bootstrap_seed <- 9

cm2_all_bootstrap <- cm2(
  fit_gllrm,
  bootstrap = TRUE,
  nsim = vignette_bootstrap_nsim,
  seed = vignette_bootstrap_seed
)
cm2_all_bootstrap
summary(cm2_all_bootstrap)$tables$aggregates
summary(cm2_all_bootstrap)
```

```{r}
cm2_subset_bootstrap <- cm2(
  fit_gllrm,
  items = diagnostic_subset_items,
  bootstrap = TRUE,
  nsim = vignette_bootstrap_nsim,
  seed = vignette_bootstrap_seed
)
cm2_subset_bootstrap
cm2_subset_bootstrap$values$bootstrap[
  c("nsim", "nused", "seed", "reestimate")
]
summary(cm2_subset_bootstrap)$tables$aggregates
summary(cm2_subset_bootstrap)
```

# CM3

CM3 uses the same default-all and explicit-subset interface as CM2. Its
aggregate contains the complete CM2 block followed by the applicable
three-way margins.

## Asymptotic CM3

The first ordinary call again omits `items`, so CM3 prepares margins for all
five fitted items.

```{r}
cm3_all_asymptotic <- cm3(fit_gllrm)
cm3_all_asymptotic
summary(cm3_all_asymptotic)$tables$aggregates
summary(cm3_all_asymptotic)
```

Selecting `P1`, `P3`, and `P5` retains the focal
`P1`-by-`P3`-by-`P5` triple and the applicable two-way and other three-way
margins, while excluding margins that require an unselected focal item.

```{r}
cm3_subset_asymptotic <- cm3(
  fit_gllrm,
  items = diagnostic_subset_items
)
cm3_subset_asymptotic
cm3_subset_asymptotic$metadata$selection$resolved_items
summary(cm3_subset_asymptotic)$tables$aggregates
summary(cm3_subset_asymptotic)
```

## Bootstrap CM3

The bootstrap comparison likewise starts with the default all-item scope and
then repeats the analysis for the subset. Both calls use the complete fitted
GLLRM, default per-sample re-estimation, and the same private Delphi-compatible
seeded stream.

```{r}
cm3_all_bootstrap <- cm3(
  fit_gllrm,
  bootstrap = TRUE,
  nsim = vignette_bootstrap_nsim,
  seed = vignette_bootstrap_seed
)
cm3_all_bootstrap
summary(cm3_all_bootstrap)$tables$aggregates
summary(cm3_all_bootstrap)
```

```{r}
cm3_subset_bootstrap <- cm3(
  fit_gllrm,
  items = diagnostic_subset_items,
  bootstrap = TRUE,
  nsim = vignette_bootstrap_nsim,
  seed = vignette_bootstrap_seed
)
cm3_subset_bootstrap
cm3_subset_bootstrap$values$bootstrap[
  c("nsim", "nused", "seed", "reestimate")
]
summary(cm3_subset_bootstrap)$tables$aggregates
summary(cm3_subset_bootstrap)
```
