---
title: "SPADI example"
author: "Andreas Kryger Jensen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SPADI example}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA)
```


Optionally install the R package
```{r, eval = FALSE}
remotes::install_github("ERRTG/gRm")
```

Then we load the package.
```{r}
library(gRm)
```

Load the data and inspect the names of the variables
```{r}
SPADI <- read.csv(url("https://erda.ku.dk/public/archives/bacc560d26b01f7a65e77a9712a92e86/SPADI.csv"))
# The source file contains one isolated P1 = 0 observation whose remaining
# record state leaves that category outside the conditional estimation sample.
# Remove that unusable category before inferring the vignette's item levels so
# every declared category has source estimation support.
SPADI <- SPADI[is.na(SPADI$P1) | SPADI$P1 > 0, , drop = FALSE]
colnames(SPADI)
```


# Defining an analysis

We start by defining a gRm analysis object based on five items and two exogenous variables from the data set. Calling the resulting object implicitly executes the print function of the object which gives basic information. For more information you can call summary on the object. Score groups are automatically defined by a median split, but they can also be specified explicitly by an argument in the construction.

```{r}
spadi_proj <- gRm(SPADI, 
                  items = sprintf("P%01d", 1:5),
                  exogenous = c("over60", "gender"))
spadi_proj
summary(spadi_proj)
```

# Rasch model

The next step is to set up a GLLRM model. We will start by fitting a standard Rasch model, so the call to gllrm is very simple. Calling the resulting object gives you basic information about the model. We see that there are neither LD nor DIF in the model. We will later see how to specify these and thus extend the Rasch model to a Graphical Log-Linear Rasch Model.

```{r}
m_rasch <- gllrm(spadi_proj)
m_rasch
```

Then we can fit the model. 

```{r}
fit_rasch <- fit(m_rasch)
fit_rasch
```

The print function used above gives a condensed summary, while the summary function gives more information. 

```{r}
summary(fit_rasch)
```

It contains two main blocks, item parameters and thresholds. The plain summary command prints both but they can also be individually printed by using the which argument.

```{r}
summary(fit_rasch, which = "parameters")
```

## Item fit

Item fit has two different views. The default is the tests view,

```{r}
item_fit(fit_rasch) #defaults to which = "tests"
```

Additional information is given by setting the view equal to items.

```{r}
item_fit(fit_rasch, which = "items")
```


## Test of global homogeneity

Test for global homogeneity follows a similar pattern where print gives the brief summary and the summary function exposes more information. The global homogeneity function will by default use the score cuts defined when constructing the analysis object, but different cuts can explicitly be defined when calling the function thus overriding the already defined ones.


```{r}
gh_rasch <- global_homogeneity(fit_rasch)
gh_rasch
```

```{r}
summary(gh_rasch)
```

## Check for local dependencies

We can check for local dependence using the local_dependence function.

```{r}
ld_rasch <- local_dependence(fit_rasch)
ld_rasch
```

```{r}
summary(ld_rasch)
```

## Check for DIF

Similarly, we can check for DIF using the dif function.

```{r}
dif_rasch <- dif(fit_rasch)
dif_rasch
```

```{r}
summary(dif_rasch)
```


# Screening

The package also implements the SCREEN J procedure from DIGRAM.

```{r}
spadi_screen <- screen(spadi_proj)
spadi_screen
```

```{r}
summary(spadi_screen)
```

The result from the screening can automatically be turned into a GLLRM model
by feeding it into `gllrm()`. The resulting model contains the LD and DIF terms
identified by SCREEN J.

```{r}
m_screen <- gllrm(spadi_screen)
m_screen
```

A visualization of the model can be obtained by the plot function.

```{r, fig.width=6, fig.height=4, out.width="100%", dpi=192}
plot(m_screen)
```

SCREEN J proposes candidate terms; the public fit boundary still verifies the
source item-category and interaction-margin support before estimation. For
these data the complete screened model has the required support and converges.

```{r}
screen_fit <- fit(m_screen)
screen_fit
```

For comparison, we also demonstrate a smaller custom GLLRM chosen from the
same candidate families.

# Setting up a general GLLRM

A custom GLLRM can also be defined by giving the specific LDs and DIFs as one-sided formula objects.

```{r}
m_gllrm <- gllrm(spadi_proj, ld = ~ P3:P4, dif = ~ P4:over60)
```

We can do a direct likelihood ratio test of the GLLRM against the Rach model

```{r}
fit_gllrm <- fit(m_gllrm)
anova(fit_rasch, fit_gllrm)
```

Global homogeneity for the fitted GLLRM includes the applicable uniform local
dependence and DIF sections.

```{r}
summary(global_homogeneity(fit_gllrm))
```


# ARI plot

The package also includes ARI plots

```{r, fig.width=6, fig.height=4, out.width="100%", dpi=192}
fit_gllrm |> ari() |> plot()
```

# M2 

```{r}
m2_fit <- m2(fit_gllrm)
m2_fit
```

```{r}
summary(m2_fit)
```


# M3

```{r}
m3_fit <- m3(fit_gllrm)
m3_fit
```

```{r}
summary(m3_fit)
```
